---
title: "FFire-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FFire-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(FFire)
```

This document contains detailed documentation and examples for the functions 
which can be found in the FFire package. In general, these function can be used 
for simple EDA and for visualizing data. The package includes the functions:

* convert_to_num()
* df_load()
* filter_data()
* plot_scatter_graph()
* plot_line_graph()

## EDA

### convert_to_num()

The function `convert_to_num()` converts columns that can be coerced to numeric
type in a data frame to a numeric type. For example, given the following 
dataframe:

```{r convert-to-num-input-data}
data_frame_c2n = data.frame(
  num_saved_as_char = c('1','2'),
  classes = c('class1', 'class2')
)
```

We can see that the column `num_saved_as_char` contains numbers which are saved
as characters! We want to tread numerical values as numerical values though.
Using `convert_to_num()` we convert the column `num_saved_as_char` to numeric
values:

```{r convert-to-num-example}
convert_to_num(data_frame_c2n, 'num_saved_as_char')
```

As we can see, the output of the functions returns the dataframe with the
column `num_saved_as_char` being formatted as numeric values!

### df_load()

This function loads and preprocesses a dataset from a CSV file,
splits the date frame into 2 parts that the first part is nice
formatted and the second part need manipulation,
correct the wrong row entries in the second part,
convert the numeric columns into correct data type,
combine the data together.

=> To be done by Stella


### filter_data()

This function calculates metrics based on a given algorithm (e.g. mean, range, 
etc.) on the used dataset (e.g. fire_train) filters the data by a given column 
name which has as specific value and by a column range. For example, given the 
following dataframe:

```{r filter-data-input-data}
   data_frame_filter_data = data.frame(
     type = c("big", "big", "small"),
     length = c(5.2, 2.1, 1.2),
     height = c(10.1, 20.1, 4.2),
     depth = c(3, 2, 1.8)
   )
  data_frame_filter_data
```

Now, let's say we want to get the mean value for the columns length and height,
while discarding the depth column and filtering by only "big" entries. We can do
this as follows:

```{r filter-data-example-1}
filter_data(data_frame_filter_data, type, "big", height, depth, mean)
```

Now, if we would like to get the range of the selected columns we can simply
run the following command:

```{r filter-data-example-2}
filter_data(data_frame_filter_data, type, "big", height, depth, range)
```

The first row shows the minimum value in the given column and the second row
shows the maximum value of the column.

## Visualization

### plot_scatter_graph()

This function can be used to plot a two dimensional scatter plot! For example, 
lets say you want to plot the following data:

```{r scatter-plot-input-data}
   data_frame_scatter_plot = data.frame(
     dryness = c(5.2, 2.1, 5.2), 
     temperature = c(10.1, 2.1, 4.2),
     Classes = c("fire", "fire", "no fire")
   )
```

We would like to visualize this data with `dryness` being on the x axis and 
`temperature` being on the y axis. We would like to label both axis with an 
uppercase name and would like to title the legend with "Presence of fire".
As we don't want the plot to be too big we want to define the height, width and
text size to ensure readability.


```{r scatter-plot-example-1,  fig.height = 5, fig.width = 5, fig.align = "center"}
 plot_scatter_graph(
   data_frame_scatter_plot,
   plot_width = 10, plot_height = 10, x_axis_data = temperature,
   y_axis_data = dryness, x_axis_label = "Temperature",
   y_axis_label = "Dryness", text_size = 20, color = Classes,
   color_label = "Presence of fire"
 )
```

Now, if we wanted to double the height of the graph we could simply change the
parameter `plot_height`:
```{r scatter-plot-example-2,  fig.height = 5, fig.width = 5, fig.align = "center"}
 plot_scatter_graph(
   data_frame_scatter_plot,
   plot_width = 10, plot_height = 20, x_axis_data = temperature,
   y_axis_data = dryness, x_axis_label = "Temperature",
   y_axis_label = "Dryness", text_size = 20, color = Classes,
   color_label = "Presence of fire"
 )
```

Or we could change the header of the legend:

```{r scatter-plot-example-3,  fig.height = 5, fig.width = 5, fig.align = "center"}
 plot_scatter_graph(
   data_frame_scatter_plot,
   plot_width = 10, plot_height = 10, x_axis_data = temperature,
   y_axis_data = dryness, x_axis_label = "Temperature",
   y_axis_label = "Dryness", text_size = 20, color = Classes,
   color_label = "Is a fire present?"
 )
```

Accordingly we can also change the labels for the x and y axis:
```{r scatter-plot-example-4,  fig.height = 5, fig.width = 5, fig.align = "center"}
 plot_scatter_graph(
   data_frame_scatter_plot,
   plot_width = 10, plot_height = 10, x_axis_data = temperature,
   y_axis_data = dryness, x_axis_label = "Factor 1",
   y_axis_label = "Factor 2", text_size = 20, color = Classes,
   color_label = "Presence of fire"
 )
```


### plot_line_graph()

With this function we can draw a line plot! Let's assume the following data
which depicts how likely a person is going to get an ice cream given the
temperature:

```{r line-plot-input-data}
   data_frame_line_plot = data.frame(
     temperature = c(5, 15, 25), 
     probability = c(0.20, 0.70, 0.95)
     )
```

We would like to visualize this data with `temperature` being on the x axis and 
`probability` being on the y axis. We would like to label both axis with an 
uppercase name. As we don't want the plot to be too big we want to define the 
height, width to ensure readability. We can do it as follows:

```{r line-plot-example-1,  fig.height = 5, fig.width = 5, fig.align = "center"}
 plot_line_graph(
   data = data_frame_line_plot,
   plot_width = 10, 
   plot_height = 10,
   x_axis_data = temperature, 
   y_axis_data = probability, 
   x_axis_label = "Temperature",
   y_axis_label = "Probability"
 )
```

Now we could change the name of the x axis:
```{r line-plot-example-2,  fig.height = 5, fig.width = 5, fig.align = "center"}
 plot_line_graph(
   data = data_frame_line_plot,
   plot_width = 10, 
   plot_height = 10,
   x_axis_data = temperature, 
   y_axis_data = probability, 
   x_axis_label = "Temperature",
   y_axis_label = "Likelihood"
 )
```
